---
title: "Bottom-up approach to create set of index"
output: html_notebook
---
## Motivation
Infinite-monkey way of identifying sets of indexes is not efficient at all. Instead, a bottom-up approach in generating a set of indexes based on a set of rules may be more appropriate. 

### Load packages
```{r}
library(stringr)
library(vwr)
library(reshape2)
library(ggplot2)       
library(viridis)
library(dplyr)
```

### Define functions 
These are the rules for creating index

```{r}
check_gc <- function(inputString){
  full.length = str_length(inputString)  
  G.count <-  str_count(inputString, "G")
  C.count <-  str_count(inputString,"C")
  return((G.count+C.count)/full.length)
}
check_repeat <- function(inputString) {
  
  possible_choice <- c("A", "T", "C", "G")
  full.length = str_length(inputString)
  chr.list = str_split(inputString, pattern = "", simplify = TRUE)
  max.counter =1
  for (base in possible_choice) {
    counter = 1
    for (i in seq(length(chr.list) - 1)) {
      if (chr.list[i] != chr.list[i + 1]) {
        next
      } else if(base == chr.list[i]){
        counter = counter + 1
        
      }
      if(counter >= max.counter){
        max.counter = counter
      }
    }

  }
  return(max.counter)
}
```

### Generate a list of 8-mer, representing 8bp indexes. 
```{r}

lenght_index <- 8
possible_choice <- c("A","T","C","G")

### for lack of a package to generate kmer, use recursive function
printAllKLength <- function(set, k){
  n = length(set)
  printAllKLengthRec(set,"",n,k)
}
printAllKLengthRec <- function(set,prefix,n,k){
  if (k == 0){
    cat(paste0(prefix,"\n"))
    return()
  }
  for (i in seq(1,n)){
    newPrefix = paste0(prefix,set[i])
    printAllKLengthRec(set,newPrefix,n,k-1)
  }
}

capture.output(printAllKLength(possible_choice,lenght_index), file = "eight_mer.txt", append = TRUE)
eight_mer <- read.table("eight_mer.txt", quote="\"", comment.char="")
eight_mer.vector <- as.character(eight_mer$V1)
```

### Create the pool 
Randomly pick a starter index. Then, pick a new index and validate it , if pass all metrics, add it to the list. Repeat until reaching the number of index.  
```{r}
library(rlist)
target.index.number <-100
GC.min <-0.3
GC.max <-0.7
runs.max <- 2
ld_threshold <- 4
output.vector =c()

### prefilter the possible option using the GC and runs
eight_mer.gc <- sapply(eight_mer.vector, check_gc)
eight_mer.repeat <- sapply(eight_mer.vector, check_repeat)
eight_mer.df <-data.frame(eight_mer.gc,eight_mer.repeat,"seq"=eight_mer.vector )
library(dplyr)
eight_mer.df %>%  filter(eight_mer.gc >GC.min) %>% filter(eight_mer.gc <GC.max) %>% filter(eight_mer.repeat <=runs.max) -> eight_mer.pass.filter.df
eight_mer.pass.filter.vector <- eight_mer.pass.filter.df$seq
names(eight_mer.pass.filter.vector) <- eight_mer.pass.filter.vector
### picking a starter, put it in a list
set.seed(92814)
starter <- sample(eight_mer.pass.filter.vector,size = 1)
output.vector<-list.append(output.vector,starter)

### start the loop
while (length(output.vector) < target.index.number){
  ### pick a new index
  new <- sample(eight_mer.pass.filter.vector,size = 1)
  ### validate the distance 
  new_ld <- levenshtein.distance(xsource = new, targets = output.vector)
  if(all(new_ld >= ld_threshold)){
    ### this is a valid index, add it to the list output.vector
    output.vector<-list.append(output.vector,new)
    
  }
}

```

### Inspect the quality of the set of indexes

```{r}
library(ggplot2)
ld_mtx <- sapply(output.vector,levenshtein.distance,targets = output.vector)
ld.df <- melt(ld_mtx)
plot.heatmap <- ggplot(ld.df,aes(as.character(Var1),as.character(Var2)))
plot.heatmap + geom_tile(aes(fill=value)) +
  theme(axis.text.x = element_blank(),axis.text.y = element_blank(),axis.title.x = element_blank(),axis.title.y = element_blank())+
  scale_fill_viridis()
hist(ld.df$value)

input.index.gc <- sapply(output.vector, check_gc)
input.index.repeat <- sapply(output.vector, check_repeat)
input.index.df <-data.frame(input.index.gc,input.index.repeat)

hist(input.index.df$input.index.gc)
summary(input.index.df$input.index.gc)
hist(input.index.df$input.index.repeat)
summary(input.index.df$input.index.repeat)
input.index.df %>% filter(input.index.repeat >=4)
```

